Gringo 2.0

### 2008-07-21 doc/documentation/paper.tex
smaller changes
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@407 3e02f596-3131-0410-8352-e04eeaec37cc

\begin{document}
\maketitle
\tableofcontents
\section{Introduction}
Answer Set Programming (ASP) \cite{baral02a} is a declarative approach for modeling and solving search problems, represented as logic programs.
An ASP system usually consists of a grounder and a solver.
First, a grounder translates a non-ground problem description into a propositional program, which can be processed by the second component of the system, the solver. This is illustrated in figure (\ref{fig:ASP}).
\begin{figure}[h]
\[
\xymatrix{*+[F]\txt{Problem\\representation}\ar[rr]^{Grounder}&&
		*+[F]\txt{Intermediate\\ representation} \ar[rr]^{Solver}&&
		*+[F]\txt{Variable\\ assignment}}
\]
\caption{Architecture of an ASP System}
\label{fig:ASP}
\end{figure}
Gringo2.0 is a state of the art and has many features that makes writing logic programs easy.
It was developed by Roland Kaminski and is orientated on the ideas of gringo1.0.
The goal of this work is to give an overview of the features supported by gringo2.0. First, the general restrictions to the input language are stated. A broad set of potentialities for writing encodings for logic programs exists. Currently, no other grounder support such a rich featured, less restricted input language.
There exist several output formats which can be used be different grounders and even a human readable format for debugging purposes.
Gringo also has an internal interface to the solver clasp \cite{gekanesc07b}, so the output can directly by parsed to the solver and answersets can be computed.
The syntax of gringo2.0 can be found in the appendix \ref{appendix}.

@@

\paragraph{--convert}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{-l}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explaned in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{-c}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answersets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore adouble minus ``--'' must be given after all gringo related parameters are set. After them die clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output al hold as clasps internal interface currently reads smodels format.
\paragraph{-p}
Parameter ``p'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo.
\paragraph{-g X}
This parameter invokes the new output format for the intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. The X can be any number from $1$ to $7$ to refer to one of the following normalforms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normalform is specified, conglomeration(7)  is assumed.

@@

\subsection{How to write a logic program}
Writing a logic program to solve a problem is rather describing a problem than describing an algorithm to solve a problem.
Therefore the problem must be encoded as a set of rules.
Typically these are divided into a problem description and an instance of the problem.
The solution to a problem can then be read from answer sets of the rules.
The procedure of writing a logic program shall now be explained with a little example.
We want to write a problem encoding for the 3-colorability problem of graph coloring.
Given an arbitraty graph, no two connected edges may have the same color.
\newline
Usually you start with the non instance domains, like what colors we do have.
\begin{verbatim}
 color(red). color(green). color(blue).
\end{verbatim}
After that we have to color the vertices, but we must give attention that each vertex only has one color.
We can do this by simply saying that a vertex is colored with one color X, and that the same vertex does not have another color Y.
\begin{verbatim}
 colored(V,C) :- not colored(V,D), C != D, color(C;D), vertex(V).
\end{verbatim}
As you can see, you have to give the domain of the variables in the rule.
The variable V is used for vertices and the variables C and D are used for colors.
After every vertex now has a color we would get all possible solutions of all vertices colorfully painted.
We now need to restrict our solutions again by saying that two vertices that are connected by an edge should not have the same color.
We can do this by using an integrity constraint.
An integrity constraint is a rule with an empty head, and if an assignment of variables fulfills the body, the assignment is not a solution of our problem.
\begin{verbatim}
 :- edge(V,U), colored(V,C), colored(U,C), vertex(V;U), color(C).
\end{verbatim}
Here we have the edge between two vertices V and U, and if both are colored with the same color C the assignment is not a solution to our problem.
\newline
We have finally written our problem description, all we need now is a problem instance. This step is fairly simple.
We just need to define several vertices and edges that do connect the vertices.
\begin{verbatim}
 vertices(1..4).
 edge(1,2). edge(1,4). edge(2,4).
 edge(3,1). edge(4,1). edge(3,2).
\end{verbatim}

If we give the problem description and instance to the grounder, it will generate a variable free version of the problem. This can then be solved by a solver and in the answer sets of the solver the colored predicate states the color of a valid coloring of all vertices.
\newline
To ease the process writing logic programs some features in the input syntax were introduced and are now explained in detail.
\subsection{Features}
\paragraph{Expanding Terms}
\subparagraph{Range Operator}
To define several atoms that stretch on one or more parameters over a certain range, two dots can used to express an increasing integer range.
Using the range operator in the head of a rule, it works as a conjunction of all the members in the range, so for every number the rule is duplicated.
\begin{verbatim}
 times(1..3, 1..2).
\end{verbatim}
will result in the facts
\begin{verbatim}
 times(1,1). times(2,1). times(3,1).
 times(1,2). times(2,2). times(3,2).
\end{verbatim}

@@

@@

\input{benchmark2.result}
\appendixname
\begin{appendix}
\label{appendix}
\input{appendix}
\end{appendix}

>>>>

\begin{document}
\maketitle
\tableofcontents
\section{Introduction}
Answer Set Programming (ASP) \cite{baral02a} is a declarative approach for modeling and solving search problems, represented as logic programs.
An ASP system usually consists of a grounder and a solver.
First, a grounder translates a non-ground problem description into a propositional program, which can be processed by the second component of the system, the solver. This is illustrated in figure (\ref{fig:ASP}).
\begin{figure}[h]
\[
\xymatrix{*+[F]\txt{Problem\\representation}\ar[rr]^{Grounder}&&
		*+[F]\txt{Intermediate\\ representation} \ar[rr]^{Solver}&&
		*+[F]\txt{Variable\\ assignment}}
\]
\caption{Architecture of an ASP System}
\label{fig:ASP}
\end{figure}
Gringo2.0 is a state of the art and has many features that makes writing logic programs easy.
It was developed by Roland Kaminski and is orientated on the ideas of gringo1.0.
The goal of this work is to give an overview of the features supported by gringo2.0. First, the general restrictions to the input language are stated. A broad set of potentialities for writing encodings for logic programs exists. Currently, no other grounder support such a rich featured, less restricted input language.
There exist several output formats which can be used be different grounders and even a human readable format for debugging purposes.
Gringo also has an internal interface to the solver clasp \cite{gekanesc07b}, so the output can directly by parsed to the solver and answersets can be computed.
Furthermore gringo can be used with clasp incrementally. Detailed information about incremental solving can be found here \cite{IncrementalASP} 
The syntax of the gringo2.0 input language can be found in the appendix \ref{appendix}.

@@

\paragraph{--convert}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{-l}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explaned in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{-c}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answersets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore adouble minus ``--'' must be given after all gringo related parameters are set. After them die clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output hold as clasps internal interface currently reads smodels format.
\paragraph{-p}
Parameter ``p'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo.
\paragraph{-g X}
This parameter invokes the new output format for the intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. The X can be any number from $1$ to $7$ to refer to one of the following normalforms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normalform is specified, conglomeration(7)  is assumed.

@@

\subsection{How to write a logic program}
Writing a logic program to solve a problem is rather describing a problem than describing an algorithm to solve a problem.
Therefore the problem must be encoded as a set of rules.
Typically these are divided into a problem description and an instance of the problem.
The solution to a problem can then be read from answer sets of the rules.
The procedure of writing a logic program shall now be explained with a little example.
We want to write a problem encoding for the 3-colorability problem of graph coloring.
Given an arbitraty graph, no two connected edges may have the same color.
\newline
Usually you start with the non instance domains, in our example the colors we use.
\begin{verbatim}
 color(red). color(green). color(blue).
\end{verbatim}
After that, we have to color the vertices, but we must give attention that each vertex has only one color.
We can do this by simply saying that a vertex is colored with one color X, and that the same vertex does not have another color Y.
\begin{verbatim}
 colored(V,C) :- not colored(V,D), C != D, color(C;D), vertex(V).
\end{verbatim}
As you can see, you have to give the domain of the variables in the rule.
The variable V is used for vertices and the variables C and D are used for colors.
After every vertex now has a color we would get all possible solutions of all vertices colorfully painted.
We now need to restrict our solutions again by saying that two vertices that are connected by an edge should not have the same color.
We can do this by using an integrity constraint.
An integrity constraint is a rule with an empty head, and if an assignment of variables fulfills the body, the assignment is not a solution of our problem.
\begin{verbatim}
 :- edge(V,U), colored(V,C), colored(U,C), vertex(V;U), color(C).
\end{verbatim}
Here we have the edge between two vertices V and U, and if both are colored with the same color C the assignment is not a solution to our problem.
\newline
We have finally written our problem description, all we need now is a problem instance. This step is fairly simple.
We just need to define several vertices and edges that do connect the vertices.
\begin{verbatim}
 vertices(1..4).
 edge(1,2). edge(1,4). edge(2,4).
 edge(3,1). edge(4,1). edge(3,2).
\end{verbatim}

If we give the problem description and instance to the grounder, it will generate a variable free version of the problem. This can then be solved by a solver and in the answer sets of the solver the ``colored'' predicate states a valid coloring of all vertices.
\newline
To ease the process writing logic programs some features in the input syntax were introduced and are now explained in detail.
\subsection{Features}
The extended features of the input syntax are now explained in detail. For a full syntax description in EBNF see appendix (\ref{appendix}).
\paragraph{Expanding Terms}
\subparagraph{Range Operator}
To define several atoms that stretch on one or more parameters over a certain range, two dots can used to express an increasing integer range.
Using the range operator in the head of a rule, it works as a conjunction of all the members in the range, so for every number the rule is duplicated.
\begin{verbatim}
 times(1..3, 1..2).
\end{verbatim}
will result in the facts
\begin{verbatim}
 times(1,1). times(2,1). times(3,1).
 times(1,2). times(2,2). times(3,2).
\end{verbatim}

@@

\section{Compiling Gringo}
For compiling gringo, a binary version of ``re2c'' is needed to compile the lexer and the parser.
TODO: What about lemon.
As the build environment ``cmake'' is used.
TODO: How to build gringo, clingo, iclingo etc...

@@

\input{benchmark2.result}
\appendixname
\begin{appendix}
\section{EBNF Syntax}
\label{appendix}
\input{appendix}
\end{appendix}
